A Study of Common Bug Fix Patterns in Rust \thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}

### 2023-11-11 response.tex

\vspace*{1em}
Thanks!

\subsection{Reorganizing related work}

\textit{\textbf{R1:} I have observed that the related work now includes several subsections focusing on APR and patch correctness prediction. It would be beneficial to reorganize and consolidate these subsections into a single coherent section. Furthermore, the section on related work should encompass studies related to bug classification in programming languages with a longer history.}

\vspace*{1em} \noindent \textbf{Response:} We have reorganized the former Sections 5.1 through 5.3 into the new Section 5.1, adding a roadmap of that subsection at its beginning.

We also added a paragraph on bug classification at the beginning of Section 5.1. It adds three new references. The first is a 1975 study conducted by Enders et al. \cite{endres1975analysis}, which examines common error patterns in the DOS/VS operating system (written in the DOS Macro Assembler langage). The second study is by Knuth~\cite{knuth89:_error_tex}, surveying errors that he added in the development of \TeX. Finally, we added a 1998 study by Flanagan et al. \cite{flanagan1998new}, where they introduce a static debugger for Scheme (a Lisp dialect), centered around the error-prone operators in the language. Additionally, we moved the initial citation of Pan et al~\cite{pan2009toward} up to this paragraph; it studies bugs in seven large-scale Java programs.

@@

\subsection{More discussion on the impact of the evolving Rust specification}

\textit{\textbf{R1:} Additionally, I would appreciate it if the authors could discuss the potential impact of the evolving Rust specification. Given that Rust is a novel programming language, it is subject to rapid grammatical changes, which in turn could influence bug characteristics and symptoms.}

\vspace*{1em} \noindent \textbf{Response:} Yes, the 2022 Rust roadmap\footnote{\url{https://blog.rust-lang.org/inside-rust/2022/04/04/lang-roadmap-2024.html}} includes future proposed changes to Rust. It is indeed a threat to external validity, and we have now discussed it in our threats to external validity (Section 4.3); it could be mitigated by refreshing the analysis results once new versions of Rust are broadly used by extant packages.

@@

\textit{Thanks a lot for the information and clarification in the rebuttal and the new paper version. I would suggest the authors make the following changes in the next paper version.}

\subsection{More discussion on the size and type of the patches}

\textit{\textbf{R2:} Inspired by the last paragraph of Section 4.1, can authors discuss which of the following is more important for Rust in the next paper version: fixing bugs and fixing compilation errors?}

\vspace*{1em} \noindent \textbf{Response:} We added a new paragraph at the end of Section 4.1. We consider bugs as potentially more critical, because an unresolved bug can remain dormant within the system and unintentionally get deployed, whereas a program with compilation errors cannot be shipped by developers.

@@

\vspace*{1em} \noindent \textit{\textbf{R2:} Please discuss why identified patches are all small in the next paper version. Basically, please add the last paragraph of Section 2.2.4 of the rebuttal to the next paper version.}

\vspace*{1em} \noindent \textbf{Response:} Added the mentioned paragraph to the Discussion section of the manuscript (third paragraph of Section 4).

@@

\textit{\textbf{R2:} In Section 2.3.2, the authors use sentences from an existing paper directly (although quotes are added). I don't think this is a good way to write an academic paper. Please paraphrase these sentences.}

\vspace*{1em} \noindent \textbf{Response:} Paraphrased the three definitions that were directly quoted.

\vspace*{1em} \noindent \textit{\textbf{R2: } I still do not understand Figure 1. I suggest the authors label ASTDiff nodes with code to ease the explanation.}

\vspace*{1em} \noindent \textbf{Response:} We have added the labels {\tt \&tail} and {\tt head.clone()} at the relevant ASTDiff nodes on the tree on the left. The nonterminals themselves don't directly correspond to code---it would be the terminals which are not present in the paths. We have also added more text to the caption of Figure 1 and point at the ExprReference in Path 1 in the text.

@@

\textit{\textbf{R2:} Line 21 on page 8: work around -> workaround}

\vspace*{1em} \noindent \textbf{Response:} Fixed.

\vspace*{1em} \noindent \textit{\textbf{R2:} Line 49 on page 8: which path is colored in purple in figure 1?}

\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer and added a reference to ``Path 3'' as in the Figure.

### 2023-11-11 discussion.tex
edit moh's changes

@@

Conversely, we were surprised to find that Ruxanne reported no patterns associated with certain Rust-related features. For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to plague many beginning Rust programmers. \verb+move+ is used to transfer the ownership of values to a closure. In some cases (e.g. multithreading), not moving the ownership could create invalid references. Rust compiler checks prevent such errors. Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many compile-time errors, but they only push their code once it is free of compile errors. We thus would not observe patterns which do not make it to the repository. While comparing the importance of bugs and compilation errors can be challenging, we tend to consider bugs as potentially more critical. This is because a bug, if left unresolved, can remain dormant within the system and inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.

@@

A threat to external validity is selection bias, where the benchmarks differ systematically from the actual world. Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects. Another threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing functionality. Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.

>>>>

While we introduce patterns that reflect changes at different locations in the code, such as \textit{BC.ref.add}, which by borrowing a variable instead of taking ownership enables users to use the variable in subsequent locations, we acknowledge that it is possible that a large patch may be harder to cluster, because it has to resemble other patches. Additionally, we believe that any clustering-based approach will likely place larger diffs further apart for the same reason that ours does: there is just more going on. Nevertheless, there is nothing inherently penalizing a larger diff. Also, we point out that the related work by~\cite{pan2009toward} specifically excludes diffs with seven statements or more, while we allow larger diffs.

@@

Conversely, we were surprised to find that Ruxanne reported no patterns associated with certain Rust-related features. For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to plague many beginning Rust programmers. \verb+move+ is used to transfer the ownership of values to a closure. In some cases (e.g. multithreading), not moving the ownership could create invalid references. Rust compiler checks prevent such errors. Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many compile-time errors, but they only push their code once it is free of compile errors. We thus would not observe patterns which do not make it to the repository.

While comparing the importance of bugs and compilation errors can be challenging, we tend to consider bugs as potentially more critical. This is because a bug, if left unresolved, can remain dormant within the system and inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.

@@

We discuss three threats to external validity: selection bias, incorrect commit/bug information reported by developers, and changes to the Rust programming language. Selection bias is an issue because our benchmarks may differ systematically from the set of Rust programs in the world. Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects. Another threat to external validity can be the developers reporting commits as bug fixing commits while they are not really fixing functionality. Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes. A third threat is that Rust is a relatively new programming language, and its syntax is prone to evolving over time. Changes in Rust's grammar, such as the introduction of new keywords like \verb+async/await+ in version 1.39, have the potential to introduce new bug categories or modify existing ones. Thus, one must consider language evolution as a potential external threat to the validity of our research findings; this threat can be mitigated by redoing our analysis in the future using our methodology with updated versions of our published artifacts.

### 2023-11-11 methodology.tex
reviewer 2 finished for now
edit moh's changes

@@

    \draw [->, green] plot [smooth, tension=1] coordinates { (2.5,1) (3,-3) (5,-5) (4,-6.5) (3.7,-8.5)};

    \draw [->, cyan] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-5) (5.5,-8.5)};

@@

    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right). A DFS will identify three paths, shown in three different colors at right.}

@@

\paragraph{Running example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff. The change (shown in the top right corner of the Figure) replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself. A DFS on this tree yields three different paths, which we show in red, green, and blue. A path represents a sequence of involved non~terminals. It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.

Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff. The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing. The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code. Instead, it is determined by the structural differences between the two ASTs of the code versions.

@@

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the blue (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description, in our opinion, is the most useful one, and we designed our weighting scheme to embed this value judgment.
We made two observations about changes. First, we saw that items (non~terminals or BC-related elements) that occur closer to AST leaves tend to be more semantically important. Second, and relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

@@

    \item refactoring: The software behavior remains unchanged after the code change. These changes are made for imporving codebase's maintainability, better readability or encapsulation.
\end{itemize}

>>>>

    \draw (0.8,-4.45) node[fill=yellow!20, minimum height=0.2cm](b0){\verb+&tail+};
    \draw (0,-5) node[](b1){ExprReference};
    \draw (0,-6) node[](b2){Path};
    \draw (0,-7) node[](b3){Ident};
    \draw (0,-8) node[](b4){tail};

    \draw (4,-4.45) node[fill=yellow!20, minimum height=0.2cm](b0){\verb+head.clone()+};
    \draw (3,-5) node[](c1){ExprMethodCall};
    \draw (3,-6) node[](c2){Path};
    \draw (3,-7) node[](c3){Ident};
    \draw (3,-8) node[](c4){head};

@@

    \draw [->, green] plot [smooth, tension=1] coordinates { (2.5,1) (3,-3) (5.2,-4.2) (4,-6.5) (3.7,-8.5)};

    \draw [->, cyan] plot [smooth, tension=1] coordinates { (3,1) (3.5,-3) (5.5,-4.2) (5.5,-8.5)};

@@

    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right). A DFS of the ASTDiff finds three paths of nonterminals (Paths 1--3), shown in three different colors at right. The first path summarizes the addition of the ExprReference {\tt \&} before tail, while the second and third paths indicate the change from a read of \texttt{head} to a method call, and a call to \texttt{clone()}, respectively.}

@@

\paragraph{Running example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff. The change (shown in the top right corner of the Figure) replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself. A DFS on this tree yields three different paths, which we label ``Path 1'' through ``Path 3'' and show in red, green, and blue. A path represents a sequence of involved non~terminals, e.g. the one starting with ExprReference in Path 1. It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.

Note that the tree depicted in Figure~\ref{fig:extraction} is not an AST; rather, it represents an ASTDiff. The paths in the ASTDiff capture the changes in the code structure between the two versions, allowing us to analyze and extract relevant information for further processing. The number of paths in the ASTDiff does not directly correspond to the number of changed elements in the source code. Instead, it is determined by the structural differences between the ASTs of the two code versions.

@@

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the blue path (Path 3) in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description, in our opinion, is the most useful one, and we designed our weighting scheme to embed this value judgment.
We made two observations about changes. First, we saw that items (non~terminals or BC-related elements) that occur closer to AST leaves tend to be more semantically important. Second, and relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

@@

    \item refactoring: The software behavior remains unchanged after the code change. These changes are made for improving the codebase's maintainability, readability, or encapsulation.
\end{itemize}

### 2023-11-11 related.tex
edit moh's changes

Automated program repair aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair. Before a repair tool modifies code, it will typically use a fault localization module to find fault locations. A fault localization module ranks program statements based on suspiciousness; one example is Tarantula by \cite{jones2005empirical}. Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\citep{naish2009spectral,xie2013theoretical}.

Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset. Following the taxonomy proposed by \cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.

\subsection{Search-Based Program Repair}
One line of research in program repair follows the competent programmer hypothesis~\citep{gopinath2014mutant}: syntactically, a faulty program is not that far away from its correct version. This hypothesis suggests the search-based program repair mindset. If the hypothesis holds, we can develop mutation operators (which change an expression or a statement) and apply them to a list of fault locations provided by the fault localization module. So, we loop through possible mutations; if the mutation of a statement does not cause all expected-successful test cases to succeed, we assume that the statement is correct and move on to the next fault location. Having a set of bug-producing patterns, like the ones we are proposing, will help researchers design targetted mutation operators that leverage domain specific insights.

Use of mutation operators in program repair became popular when \cite{forrest2009genetic} and \cite{nguyen2009using} developed GenProg, a program repair tool that uses genetic programming. The main idea behind this tool is that statements follow certain patterns in a codebase. Therefore, if we find the correct version of a faulty statement somewhere in the program, we can use that version in place of the faulty original. This tool showed promising results as it managed to find patches without any additional annotations or human involvements.

However, \cite{arcuri2011practical} observed that GenProg mostly found the patch while carrying out random initialization---that is, before GenProg's evolution begins. Their tools TrpAutoRepair~\citep{qi2013efficient} and RSRepair~\citep{qi2014strength} advocate that test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.

@@

\subsection{Pattern-based Program Repair and Mining Bug Fix Patterns}

The main intuition in pattern-based program repair methods is that bug fixes follow certain patterns. Thus, having a collection of common bug fix patterns is a crucial step for developing such repair tools. \cite{pan2009toward} analyzed seven large-scale widely used Java projects and obtained 27 common bug fix patterns. They used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements. They ignored larger bug fix hunks as they tended to exhibit random changes and did not have observable meaningful patterns. Though they did the analysis on Java projects, their reported patterns were not specific to Java. \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method. In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.

@@

We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining}. Like~\cite{hanam2016discovering}, we used DBSCAN to cluster fixes. Moreover, similar to~\cite{yang2022mining}, we introduced patterns in two different categories. In our case there were general and language-specific patterns; for Rust, the language-specific patterns were related to the borrow checker (BC patterns). However, unlike our paper, \cite{yang2022mining} needed to do a much more elaborate manual analysis, as they had only collected general information about the single hunk changes (e.g. the number of variables, or arguments).

\subsection{Automated Patch Correctness Assessment}

Overfitting poses a substantial challenge for automated program repair (APR) systems, where synthesized patches may pass test cases but fail to be correct. Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses. While our project has a different objective than APCA, we share similarities in design decisions with APCA systems, particularly in feature extraction. \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects. ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks. Similarly, our feature extraction involves manual crafting of features. In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages. Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the changed part and the unchanged part (context code) of patches. Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of systems that attempt to automatically assess patch correctness (which is not our goal).

@@

\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that utilizes log messages accompanying code changes to learn representations capturing semantic information. By addressing challenges in distinguishing between added and removed code and incorporating an attention mechanism, CC2Vec outperforms previous techniques for code change representation. The model employs a hierarchical attention network, which incorporates a multi-level bidirectional GRU recurrent neural network, to encode information about changed code tokens, lines, and hunks into vectors representing added and removed changes. These vectors are then combined and fed into a hidden layer for predicting the target function. Evaluation across log message generation, bug fixing patch identification, and just-in-time defect prediction tasks demonstrates the superior performance of CC2Vec compared to state-of-the-art approaches in all three tasks.

Our work tends to be more similar to that of~\cite{hanam2016discovering}. However, our novel program embedding approach differs from what they did on certain key decisions. In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements. Embedding programs based on this table resulted in an order-sensitive embedding. However, it made the datapoints sparse, and our datapoints are more dense than theirs. More specifically, the number of our dimensions is more than 11$\times$ smaller.

\subsection{Refactoring tools for Rust}
Clippy is widely known as the best code refactoring tool for Rust. It can identify over 600 common mistakes and offer automated fixes. It provides users with the flexibility to adjust the linting level, allowing them to control the number of lint suggestions they receive. As discussed in Section~\ref{sec:bc_patterns}, we found that Clippy can detect and address five of our borrow-checker-related patterns. \cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations, including renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters. In a more advanced use case, \cite{ling2022rust} proposed CRustS, a tool designed for refactoring C2Rust output. C2Rust is a C to Rust transpiler that converts C code into Rust syntax but retains the unsafe semantics of C. By relaxing the constraint of preserving semantics during transformations, CRustS increases the proportion of transformed code that successfully passes the safety checks of the Rust compiler.

@@

\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours. They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the Rust standard library. They analyzed the motivation behind unsafe code usage and removal, in addition to recording 70 memory-safety issues and 100 concurrency bugs. They also provided Rust programmers with some suggestions and insights to develop better Rust programs. Using the results of their manual analysis, they designed two bug detectors, and provided recommendations for developing more bug detectors in the future. Our automated approach to finding bug patterns has similar implications to their work but can operate at a far larger scale.

\cite{li2021mirchecker} present MirChecker, a fully automated bug detection framework for Rust. This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR). The tool exploits the insights obtained from manually observing existing bugs (by studying reported CVEs) in Rust code bases. Using both numerical (e.g. integer bounds) and symbolic (e.g. modelling memory) information, the framework detects errors by using constraint solving techniques. MirChecker detected 33 new bugs, including 16 memory safety faults, across 12 Rust crates.

>>>>

Automated program repair (APR) aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair. Before a repair tool modifies code, it will typically use a fault localization module to find fault locations. A fault localization module ranks program statements based on suspiciousness; one example is Tarantula by \cite{jones2005empirical}. Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\citep{naish2009spectral,xie2013theoretical}.

\subsection{Automated Program Repair}
We next discuss a number of approaches for automated program repair. First, however, we discuss previous work that aimed to classify bugs---program repair must have a model of what it is attempting to repair in the first place. We then present two mindsets for program repair: search-based and pattern-based. Finally, we survey related work in automated patch correctness assessment, which aims to ensure that the generated changes are valid.

\paragraph{Bug Classification.}
A foundational bug study dates back to~\cite{endres1975analysis}, which classified the bugs that were fixed in a particular release of IBM's DOS/VS operating system, written in the DOS Macro Assembler language. Endres contended that categorizing bugs can aid in uncovering error causes, fixing them, and preventing their recurrence. The Group B of bugs in that work include, among others, bugs that could be prevented by the use of a different programming language, e.g. ``assignment, loading, or saving of address registers forgotten,'' which is somewhat similar to our Rust-specific bug patterns. Later on, \cite{knuth89:_error_tex} analyzed the errors that he made in the development of \TeX, written in the WEB literate programming language. This was perhaps one of the first works in this domain where there is a publicly-available artifact (the \TeX~source code) supporting the data (i.e. the bugs identified), unlike the proprietary DOS/VS code. Knuth includes 15 categories of errors; the most relevant ones to this work are ``language liabilities,'' ``mismatches between modules,'' ``reinforcements of robustness,'' and ``surprising scenarios.'' \cite{flanagan1998new} described MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp). MrSpidey is a front-end to a ESC-style static verification engine that aims to detect bug patterns where operators are passed invalid arguments. The engine relies on program invariants and presents information about attempted proofs of the invariants to the developer. Most recently, \cite{pan2009toward} analyzed seven large-scale widely used Java projects and manually extracted 27 common bug fix patterns.

Moving from bug classification to our application, program repair, we follow the taxonomy of~\cite{liu2018survey} and propose two mindsets: search-based program repair and pattern-based program repair. Our work applies to both mindsets, but is most directly applicable to pattern-based program repair.

\paragraph{Search-Based Program Repair.}
One line of research in program repair follows the ``competent programmer'' hypothesis~\citep{gopinath2014mutant}: syntactically, a faulty program is not that far away from its correct version. This hypothesis suggests the search-based program repair mindset. If the hypothesis holds, we can develop mutation operators (which change an expression or a statement) and apply them to a list of fault locations provided by the fault localization module. So, we loop through possible mutations; if the mutation of a statement does not cause all expected-successful test cases to succeed, we assume that the statement is correct and move on to the next fault location. Having a set of bug-producing patterns, like the ones we are proposing, will help researchers design targetted mutation operators that leverage domain specific insights.

Use of mutation operators in program repair became popular with GenProg~(\cite{forrest2009genetic,nguyen2009using}), a program repair tool that uses genetic programming. The main idea behind this tool is that statements follow certain patterns in a codebase. Therefore, if we find the correct version of a faulty statement somewhere in the program, we can use that version in place of the faulty original. This tool showed promising results as it managed to find patches without any additional annotations or human involvements.

However, \cite{arcuri2011practical} observed that GenProg mostly found patches while carrying out random initialization---that is, before GenProg's evolution begins. Their tools TrpAutoRepair~\citep{qi2013efficient} and RSRepair~\citep{qi2014strength} advocate that test case prioritization is necessary while using genetic programming, as fitness evaluation is an expensive part of the repair pipeline.

@@

\paragraph{Pattern-based Program Repair and Mining Bug Fix Patterns.}
The main intuition in pattern-based program repair methods is that bug fixes follow certain patterns. Thus, having a collection of common bug fix patterns is a crucial step for developing such repair tools. For this mindset, the search is for a program fix (i.e. a delta) rather than for a correct program.

As previously cited, \cite{pan2009toward}, used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements. They ignored larger bug fix hunks as such hunks tended to exhibit random changes and did not have observable meaningful patterns. Though they did the analysis on Java projects, their reported patterns were not specific to Java. \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method. In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.

@@

We would characterize our work as the Rust version of what was done in~\cite{hanam2016discovering} and~\cite{yang2022mining} for JavaScript and Python, respectively. Like~\cite{hanam2016discovering}, we used DBSCAN to cluster fixes. Moreover, similar to~\cite{yang2022mining}, we introduced patterns in two different categories. In our case, there were general and language-specific patterns. The Rust language-specific patterns were related to the borrow checker (BC patterns). However, unlike our paper, \cite{yang2022mining} needed to do a much more elaborate manual analysis, as they had only collected general information about the single hunk changes (e.g. the number of variables, or arguments).

\paragraph{Automated Patch Correctness Assessment.}
The final step in automated program repair is to validate the proposed changes. One substantial challenge is overfitting, where synthesized patches may pass test cases but fail to be correct. Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses. While our project has a different objective than APCA, we share similarities in design decisions with APCA systems, particularly in feature extraction. \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects. ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks. Similarly, our feature extraction involves manual crafting of features. In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages. Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the changed part and the unchanged part (context code) of patches. Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of systems that attempt to automatically assess patch correctness (which is not our goal).

@@

\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that uses log messages accompanying code changes to learn representations capturing semantic information. By addressing challenges in distinguishing between added and removed code and incorporating an attention mechanism, CC2Vec outperforms previous techniques for code change representation. The model employs a hierarchical attention network, which incorporates a multi-level bidirectional GRU recurrent neural network, to encode information about changed code tokens, lines, and hunks into vectors representing added and removed changes. These vectors are then combined and fed into a hidden layer for predicting the target function. Evaluation across log message generation, bug fixing patch identification, and just-in-time defect prediction tasks demonstrates the superior performance of CC2Vec compared to the previous state-of-the-art approach in all three tasks.

Our work is more like that of~\cite{hanam2016discovering}. However, our novel program embedding approach differs from theirs on certain key decisions. In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements. Embedding programs based on this table resulted in an order-sensitive embedding. However, it made the datapoints sparse; our datapoints are more dense than theirs. More specifically, the number of our dimensions is more than 11$\times$ smaller.

\subsection{Refactoring tools for Rust}
Clippy is widely known as the best code refactoring tool for Rust. It can identify over 600 common mistakes and offer automated fixes.
% true but doesn't add anything to our point
%It provides users with the flexibility to adjust the linting level, allowing them to control the number of lint suggestions they receive.
As discussed in Section~\ref{sec:bc_patterns}, we found that Clippy can detect and address five of our borrow-checker-related patterns. Beyond Clippy, \cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations, including renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters. In a more advanced use case, \cite{ling2022rust} proposed CRustS, a tool designed for refactoring C2Rust output. C2Rust is a C to Rust transpiler that converts C code into Rust syntax but retains the unsafe semantics of C. By relaxing the constraint of preserving semantics during transformations, CRustS increases the proportion of transformed code that successfully passes the safety checks of the Rust compiler.

@@

\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours. They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the Rust standard library. They analyzed the motivation behind unsafe code usage and removal, in addition to recording 70 memory-safety issues and 100 concurrency bugs. They also provided Rust programmers with some suggestions and insights to develop better Rust programs. Using the results of their manual analysis, they designed two bug detectors, and provided recommendations for developing more bug detectors in the future. Our approach to finding bug patterns has similar implications to their work but, due to automation, can operate at a far larger scale.

\cite{li2021mirchecker} present MirChecker, a fully automated bug detection framework for Rust. This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR). The tool exploits the insights obtained from manually observing existing bugs (by studying reported CVEs) in Rust code bases. Using custom abstract domains that consider both numerical (e.g. integer bounds) and symbolic (e.g. modelling memory) information, the framework detects errors using constraint solving techniques. MirChecker detected 33 new bugs, including 16 memory safety faults, across 12 Rust crates. While MirChecker automatically detects bugs, the bug classes that it detects must be manually programmed into the system, and our work is complementary to MirChecker in potentially proposing novel classes of bugs that need to be fixed.
